---
title: "Simulation Testing Data Limited Methods **FLife**"
subtitle: "Catch Only"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
github_document:
  mathjax: TRUE
pdf_document:
  fig_width: 6 
  fig_height: 4 
tags: [FLR]
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
---
 
```{r, ini, echo=FALSE, results='hide', message=FALSE, warnings=FALSE, cache=FALSE,eval=!TRUE}
library(knitr)
source("R/ini.R")
```

```{r knitr_init, echo=FALSE, results="hide", eval=!FALSE}
library(knitr)
## Global options
opts_chunk$set(cache     =TRUE,
               echo      =FALSE,
               eval      =TRUE,
               prompt    =FALSE,
               comment   =NA,
               message   =FALSE,
               warning   =FALSE,
               tidy      =TRUE,
               fig.height=6,
               fig.width =8,
               fig.path  ="tex/simtest/srar-",
               cache.path="cache/simtest/sra/")
```

```{r}
options(digits=3)

iFig=0
```

```{r, methods_to_go_in_pkgs, echo=FALSE,eval=FALSE}
#source('~/Desktop/sea++/mydas/pkg/R/genTime.R')
#source('~/Desktop/sea++/mydas/pkg/R/mlz.R')
#source('~/Desktop/sea++/mydas/pkg/R/oemLn.R')
#source('~/Desktop/sea++/mydas/pkg/R/mlz.R')
#source('~/Desktop/sea++/mydas/pkg/R/omOut.R')
#source('~/Desktop/sea++/mydas/pkg/R/lbspr.R')
#source('~/Desktop/flr/mpb/R/biodyn-sra.R')
```


# Introduction

This tutorial describes how to simuation test data limited methods in `FLR` using a variety of other packages.

## Required packages

To follow this tutorial you should have installed the following packages:

- FLR: [FLCore](http://www.flr-project.org/FLCore/), [FLBRP](http://www.flr-project.org/FLBRP/), [FLasher](http://www.flr-project.org/FLash/), [FLife](http://www.flr-project.org/FLife/)

for example

```{r, install, eval=FALSE}
install.packages(c("FLCore"),  repos="http://flr-project.org/R")
install.packages(c("FLBRP"),   repos="http://flr-project.org/R")
install.packages(c("FLasher"), repos="http://flr-project.org/R")
install.packages(c("FLife"),   repos="http://flr-project.org/R")
```

```{r, pkgs}
# Load  packages
library(ggplot2)
library(plyr)
library(reshape)

library(FLCore)
library(ggplotFL)
library(FLBRP)
library(FLasher)
library(FLife)
```

# Operating Model

Turbot

```{r}
#source('~/Desktop/sea++/mydas/pkg/R/genTime.R')

lh=FLPar(c(linf= 59.1,  k=0.28, t0=-0.4, a=0.01111,b=3.15,a50=4.0, l50=43.25),units="NA")
lh=lhPar(lh)
eq=lhEql(lh)

gTime=c(round(gt(eq)))

fbar(eq)=refpts(eq)["msy","harvest"]%*%FLQuant(c(rep(.1,19),
                                              seq(.1,2,length.out=30),
                                              seq(2,1.0,length.out=gTime)[-1],
                                              rep(1.0,61)))[,1:105]
om=as(eq,"FLStock")
om=fwd(om,f=fbar(om)[,-1], sr=eq)
```

```{r turbot-ts-ref}
plot(FLQuants(om, 
          "f" =   function(x) fbar(x)%/%refpts(eq)["msy","harvest"], 
          "ssb" = function(x) ssb(x)%/%refpts( eq)["msy","ssb"], 
          "catch"=function(x) landings(x)%/%refpts(eq)["msy","yield"],
          "rec" = function(x) rec(x)%/%refpts( eq)["msy","rec"])) + 
  geom_hline(aes(yintercept=1),col="red",linetype=2)+
  theme_bw() 
```

**Figure `r iFig=iFig+1; iFig`** Time series relative to MSY benchmarks.

# Catch Only Based Methods


```{r}
## MP
library(mpb)
#source('~/Desktop/flr/mpb/R/setMP.R')
#source('~/Desktop/flr/mpb/R/pellam.R')

setMP=mpb:::setMP
p=mpb:::p

load("/home/laurence/Desktop/sea++/mydas/pkg/data/turbot.RData")

mp=setMP(as(window(om,end=55),"biodyn"),
         r =median(prior["r"],na.rm=T),
         k =median(prior["v"],na.rm=T),
         b0=0.8,
         p =median(p(prior["bmsy"]/prior["v"]),na.rm=TRUE))

nits=dims(om)$iter
set.seed(1234)
 srDev=rlnoise(nits,FLQuant(0,dimnames=list(year=1:100)),0.3,b=0.0)

eq=FLCore::iter(eq,seq(nits))

mp=setMP(as(window(om,end=55),"biodyn"),
                 r =median(prior["r"],na.rm=T),
                 k =median(prior["v"],na.rm=T),
                 b0=0.8,
                 p =median(p(prior["bmsy"]/prior["v"]),na.rm=TRUE))
```

```{r, eval=FALSE}
res=mseSRA(om,eq,mp,
              start=mseStart[stock],end=mseStart[stock]+40,
              ftar=ftar,btrig=btrig,
              srDev=srDev)
```



# References

# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Or send a pull request to <https://github.com/flr/doc/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage, <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* FLasher: `r packageVersion('FLasher')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author information

**Laurence KELL**. laurie@seaplusplus.co.uk


## Acknowledgements

This vignette and the methods documented in it were developed under the MyDas project funded by the Irish exchequer and EMFF 2014-2020. The overall aim of MyDas is to develop and test a range of assessment models and methods to establish Maximum Sustainable Yield (MSY) reference points (or proxy MSY reference points) across the spectrum of data-limited stocks.
