---
title: "Simulation Testing Data Limited Methods **FLife**"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
github_document:
  mathjax: TRUE
pdf_document:
  fig_width: 6 
  fig_height: 4 
tags: [FLR]
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
---
 
```{r, ini, echo=FALSE, results='hide', message=FALSE, warnings=FALSE, cache=FALSE,eval=!TRUE}
library(knitr)
source("R/ini.R")
```

```{r knitr_init, echo=FALSE, results="hide", eval=!FALSE}
library(knitr)  
## Global options
opts_chunk$set(cache     =TRUE,
               echo      =FALSE,
               eval      =TRUE,
               prompt    =FALSE,
               comment   =NA,
               message   =FALSE,
               warning   =FALSE,
               tidy      =TRUE,
               fig.height=6,
               fig.width =8,
               fig.path  ="tex/simtest/len-",
               cache.path="cache/simtest/len/")
```

```{r}
options(digits=3)

iFig=0
```


# Introduction

This tutorial describes how to simuation test data limited methods in `FLR` using a variety of other packages.

## Required packages

To follow this tutorial you should have installed the following packages:

- FLR: [FLCore](http://www.flr-project.org/FLCore/), [FLBRP](http://www.flr-project.org/FLBRP/), [FLasher](http://www.flr-project.org/FLash/), [FLife](http://www.flr-project.org/FLife/)

for example

```{r, install, eval=FALSE}
install.packages(c("FLCore"),  repos="http://flr-project.org/R")
install.packages(c("FLBRP"),   repos="http://flr-project.org/R")
install.packages(c("FLasher"), repos="http://flr-project.org/R")
install.packages(c("FLife"),   repos="http://flr-project.org/R")
```

```{r, pkgs}
# Load  packages
library(ggplot2)
library(plyr)
library(reshape)

library(popbio)

library(FLCore)
library(ggplotFL)
library(FLBRP)
library(FLasher)
library(FLife)
library(mydas)
```

# Operating Model

Turbot

```{r, om}
lh=FLPar(c(linf= 59.1,  k=0.28, t0=-0.4, a=0.01111,b=3.15,a50=4.0, l50=43.25),units="NA")
lh=lhPar(lh)
eq=lhEql(lh)

gTime=c(round(gt(eq)))
fbar(eq)=refpts(eq)["msy","harvest"]%*%FLQuant(c(rep(.1,19),
                                              seq(.1,2,length.out=30)[-30],
                                              seq(2,1.0,length.out=gTime)[-1],
                                              rep(1.0,61)))[,1:105]

om=as(eq,"FLStock")
om=fwd(om,f=fbar(om)[,-1], sr=eq)
```

```{r, ts}
plot(FLQuants(om, 
          "f" =   function(x) fbar(x)%/%refpts(eq)["msy","harvest"], 
          "ssb" = function(x) ssb(x)%/%refpts( eq)["msy","ssb"], 
          "catch"=function(x) landings(x)%/%refpts(eq)["msy","yield"],
          "rec" = function(x) rec(x)%/%refpts( eq)["msy","rec"])) + 
  geom_hline(aes(yintercept=1),col="red",linetype=2)+
  theme_bw() 
```

**Figure `r iFig=iFig+1; iFig`** Time series relative to MSY benchmarks.

# Length Based Methods


Based on Beverton and Holt  $L_{F} = \frac{L\infty +\frac{F+M}{K}L_c}{1+\frac{F+M}{K}}$

+ $L_{current}/L_{F=M}$ 
+ $M/(Z_{current}-M)$
+ $F_{0.1}/(Z_{current}-M)$
+ $LBSPR$ Length-based spawning potential ratio (Hordyk et al. 2015) 
+ $Lime$ (Rudd and Thorson, 2018) mixed effects non-equilibrium



## MLZ

[MLZ](https://cran.r-project.org/web/packages/MLZ/index.html) is a package that facilitates data preparation and estimation of mortality with statistical diagnostics using the mean length-based mortality estimator and several extensions.


```{r, len}
ts   =omSmry(om,eq,lh)
mnLen=as.FLQuant(with(ts,data.frame(data=cln,year=year,iter=iter)))

plot(mnLen)
```
**Figure `r iFig=iFig+1; iFig`** Mean length of catch turbot. 

```{r, priors}
#growth<-vonB
prior=priors(lh)
```

```{r, mlz}
library(MLZ)

res=mydas:::mlz(mnLen[,ac(40:60)],prior)

res
```


## LB-SPR

[LBSPR](https://cran.r-project.org/web/packages/LBSPR/vignettes/LBSPR.html)
is a R package for simulation and estimation using life-history ratios and length composition data



```{r, lbspr}
library(LBSPR)
```

```{r, alk}
ak=alk(lh)  
```

```{r, lfd}
lfd=lenSample(catch.n(om)[,20:65],ak,nsample=500)
```

```{r, oem}
ggplot(melt(lfd[,seq(1,45,10)]))+
  geom_histogram(aes(length,weight=value),binwidth=1)+
  facet_grid(year~iter,scale="free")+
  xlab("Length (cm)")+ylab("Frequency")+
  coord_cartesian(xlim=c(0,mean(lh["linf"])))
```
**Figure `r iFig=iFig+1; iFig`** Observation error model for turbot. 


```{r, sa}
library(LBSPR)
library(popbio)

prior=priors(lh)

lb=lbspr(lfd,prior)  
```


```{r, spr}
ggplot(melt(sweep(lb["SPR"],c(1,3),lb["SPR","40"],"/")))+
  geom_boxplot(aes(ac(year),value))+
  scale_x_discrete(breaks=seq(20,60,10))+
  ylab("SPR")+xlab("Year")+theme_bw()
```
**Figure `r iFig=iFig+1; iFig`** Estimates of SPR for turbot. 

```{r, fm}
ggplot(melt(sweep(lb["FM"],c(1,3),lb["FM","40"],"/")))+
  geom_boxplot(aes(ac(year),value))+
  scale_x_discrete(breaks=seq(20,60,10))+
  ylab("F")+xlab("Year")+theme_bw()
```
**Figure `r iFig=iFig+1; iFig`** Estimates of $F/M$ for turbot. 


# References

# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Or send a pull request to <https://github.com/flr/doc/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage, <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* FLasher: `r packageVersion('FLasher')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author information

**Laurence KELL**. laurie@seaplusplus.co.uk


## Acknowledgements

This vignette and the methods documented in it were developed under the MyDas project funded by the Irish exchequer and EMFF 2014-2020. The overall aim of MyDas is to develop and test a range of assessment models and methods to establish Maximum Sustainable Yield (MSY) reference points (or proxy MSY reference points) across the spectrum of data-limited stocks.
