---
title: "**mydas**"
subtitle: "Conditioning Operating Models on Data"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
github_document:
  mathjax: TRUE
pdf_document:
  fig_width: 6 
  fig_height: 4 
tags: [FLR]
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
---

http://rpubs.com/pl202/436164

```{r knitr_init, echo=FALSE, results="hide"}
library(knitr)
## Global options
opts_chunk$set(cache     =TRUE,
               cache.path='cache/conditioning/',
               echo      =TRUE,
               eval      =TRUE,
               prompt    =FALSE,
               comment   =NA,
               message   =FALSE,
               warning   =FALSE,
               tidy      =FALSE,
               fig.height=6,
               fig.width =8,
               fig.path  ='tex/conditioning-')


iFig=0
```


```{r, pkgs, message=FALSE}
library(ggplotFL)
library(plyr)
library(reshape)
library(FLBRP)
library(FLife)
```

```{r, theme, echo=FALSE}
theme_set(theme_bw())
options(digits=3)
```

## Install FLR packages

To follow this tutorial you will have to install a number of packages, either from CRAN or from [www.flr-project.org](http://www.flr-project.org) where variety of packages and [tutorials](https://www.flr-project.org/doc/) are available.

Install FLR from [https://www.flr-project.org/](www.flr-project.org/)
```{r, eval=FALSE}
install.packages(c("FLCore","FLFishery","FLasher","FLBRP","mpb","FLife"), 
             repos="http://flr-project.org/R")
```

`devtools` needs to be installed and then loaded so that the `mydas` package can be installed from this GitHub repository.

```{r}
install.packages("devtools",dependencies=TRUE)
```

```{r, echo=TRUE, eval=FALSE}
library(devtools)

devtools::install_github("lauriekell/mydas-pkg")
```

## Load Libraries
```{r}
library(plyr)
library(reshape)
library(ggplot2)
```

```{r}
library(FLCore)
library(FLasher)
library(FLBRP)
library(FLife)
library(mydas)
```


```{r}
```


**Figure `r iFig=iFig+1; iFig`** 

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* FLPKG: `r # packageVersion('FLPKG')`
* **Compiled**: `r date()`
* **Git Hash**: `r system("git log --pretty=format:'%h' -n 1", intern=TRUE)`

## Author information

**Laurence KELL**. laurie@seaplusplus.co.uk

## Acknowledgements


# References {#References}


