---
title: "mydas"
subtitle: "Life History Relationships"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "Laurence Kell"
output: rmarkdown:::pdf_document
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{FLife-WKLife}
  %\VignetteEncoding{UTF-8}
bibliography: refs.bib
tags: FLPKG FLR
license: Creative Commons Attribution-ShareAlike 4.0 International
---


# mydas package

```{r knitr_init, echo=FALSE, results="hide"}
library(knitr)
## Global options
opts_chunk$set(cache     =!TRUE,
               echo      =TRUE,
               eval      =TRUE,
               prompt    =FALSE,
               comment   =NA,
               message   =FALSE,
               warning   =FALSE,
               tidy      =FALSE,
               fig.height=6,
               fig.width =8)

iFig=0
```


```{r, pkgs, message=FALSE}
library(mydas)
```

```{r, theme, echo=FALSE}
theme_set(theme_bw())
options(digits=3)
```


## Install FLR packages

To follow this tutorial a number of packages need to be installed, either from CRAN or from [www.flr-project.org](http://www.flr-project.org) where variety of packages and [tutorials](https://www.flr-project.org/doc/) are available.

```{r, eval=FALSE}
install.packages(c("FLCore","FLFishery","FLasher","FLBRP","mpb","FLife"), 
             repos="http://flr-project.org/R")
```

`devtools` also needs to be installed and loaded so that the `mydas` package can be installed from the GitHub repository.

```{r}
install.packages("devtools",dependencies=TRUE)
```

```{r, echo=TRUE, eval=FALSE}
library(devtools)

devtools::install_github("lauriekell/mydas-pkg")
```

## Load Libraries
```{r}
library(FLCore)
library(FLasher)
library(FLBRP)
library(FLife)
library(mydas)
```


### Life history parameters
Get fishbase data
```{r}
stocks=data.frame(
   species=c("Psetta maxima","Scophthalmus rhombus","Raja clavata",
              "Pollachius pollachius","Sprattus sprattus sprattus"),
   name=   c("Turbot",       "Brill",               "Ray",        
             "Pollack",      "Sprat"))

load(url("https://github.com//fishnets//fishnets//blob//master//data//fishbase-web//fishbase-web.RData?raw=True"))

finfish=subset(fb,as.character(species)%in%as.character(stock$species))
```

```{r}


names(lh)[c(14,17)] = c("l50","a50")
lh=lh[,c("linf","k","t0","a","b","a50","l50")]

head(lh)
```

# Life history parameters
```{r, data}
data(mydas)
```


```{r, fig.height=8, echo=FALSE}
library(GGally)

my_smooth <- function(data,mapping,...){
  ggplot(data=data,mapping=mapping)+
  geom_point(...,size=.5)+
  geom_smooth(...,method="lm",se=FALSE)}

my_density <- function(data,mapping,...){
  ggplot(data=data,mapping=mapping)+
  geom_density(...,lwd=1)}

ggpairs(transform(wklife[,-c(1:5)],linf=log(linf),k=log(k),l50=log(l50)),
  lower = list(continuous = wrap(my_smooth)),
  diag=list(continuous=wrap(my_density,alpha=0.2)),
  title = "")+
  theme(legend.position ="none",
  panel.grid.major =element_blank(),
  axis.ticks       =element_blank(),
  axis.text.x      =element_blank(),
  axis.text.y      =element_blank(),
  panel.border     =element_rect(linetype = 1, colour="black", fill=NA))
```

**Figure `r iFig=iFig+1; iFig`** Pairwise scatter plots of life history parameters.


# Equilibrium Dynamics

Create an `FLPar`
```{r FLPar}
wkpar=as(wklife[,6:13],"FLPar")
attributes(wkpar)[names(wklife)[1:5]]=wklife[,1:5]
```

Then use life history relationships to estimate missing values  
```{r m-gislason}
par <- lhPar(wkpar)
units(par) <- c("", "", "cm", "", "cm", "cm", "cm", "cm")
```

and then to derive vectors for processses such as natural mortality 
 
```{r eqls}
eql=lhEql(par)
```

```{r vectors}
sel<-function(x) 
  catch.sel(x)%/%fapex(catch.sel(x))

ggplot(FLQuants(eql,"m","catch.sel"=sel,"mat","catch.wt"))+
  geom_line(aes(age,data,col=attributes(wkpar)$name[iter]))+
  facet_wrap(~qname,scale="free")+
  scale_x_continuous(limits=c(0,15))+ 
  guides(colour=guide_legend(title="Species",title.position="top"))
```

**Figure `r iFig=iFig+1; iFig`** Vectors of m, selection pattern, maturity and weight-at-age.

and estimate equilibrium dynamics and reference points, e.g. for lemon sole

```{r eql,eval=FALSE}
plot(iter(eql,3))
```

**Figure `r iFig=iFig+1; iFig`** Equilibrium curves for ling.

# Simulation

Create a forward projection, i.e. an `FLStock` from an equilibrium object

```{r eql-lmsl}
lmsl=as(iter(eql,7),"FLStock")

units(mat(lmsl))="NA"
units(harvest(lmsl))="f"
plot(lmsl)
```

**Figure `r iFig=iFig+1; iFig`** Simulate a stock with increasing F

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* FLPKG: `r # packageVersion('FLPKG')`
* **Compiled**: `r date()`
* **Git Hash**: `r system("git log --pretty=format:'%h' -n 1", intern=TRUE)`

## Author information

**Laurence KELL**. laurie.kell.es


## Acknowledgements

This vignette and many of the methods documented in it were developed under the MyDas project funded by the Irish exchequer and EMFF 2014-2020. The overall aim of MyDas is to develop and test a range of assessment models and methods to establish Maximum Sustainable Yield (MSY) reference points (or proxy MSY reference points) across the spectrum of data-limited stocks.

# References {#References}


