---
title: "MyDas"
subtitle: "Observation Error Model for Data Limited Stocks"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
github_document:
  mathjax: TRUE
pdf_document:
  fig_width: 6 
  fig_height: 4 
tags: [FLR]
bibliography: /home/laurence/Desktop/refs.bib
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
---

```{r, ini, echo=FALSE, results='hide', message=FALSE, warnings=FALSE, cache=FALSE,eval=!TRUE}
library(knitr)
#source("R/ini.R")
```
 
```{r knitr_init, echo=FALSE, results="hide",eval=TRUE}
library(knitr)
## Global options
opts_chunk$set(echo    =!TRUE,
               eval    =TRUE,
               cache   =TRUE,
               cache.path="cache/oem/",
               prompt  =FALSE,
               comment =NA,
               message =FALSE,
               tidy    =FALSE,
               warnings=FALSE,
               fig.height=4.5,
               fig.width =6,
               fig.path  ="tex/oem-")
iFig=0
```


[](#top)

[Introduction](#Introduction)

[Installation](#Installation)

[Examples](#Examples)

[More information](#More information)

[References](#References)


# Introduction {#Introduction}

In Management Strategy Evaluation (MSE) an Operating Model (OM) is used to simulate resource dynamics to evaluate the performance of a Management Procedure (MP). Where the MP is the combination of pre-defined data, together with an algorithm to which such data are input to provide a value for a management control measure. 

The link between the OM and the MP is the Observation Error Model (OEM), which generates fishery-dependent or independent resource monitoring data. The OEM models the uncertainties due to sampling and limited data. 


[Back to Top](#top)

# Installation

To run the code in this vignette a number of packages need to be installed, both from CRAN and 
[FLR](http://www.flr-project.org), where  [tutorials](https://www.flr-project.org/doc/) are available.


## Load Libraries

### CRAN
All the examples make extensive use of the packages of [Hadley Wickham](http://hadley.nz/)

```{r}
library(plyr)
library(reshape)
library(ggplot2)
```

### FLR

The FLR packages can be installed from  [www.flr-project.org](http://www.flr-project.org)

```{r, eval=FALSE}
install.packages(c("FLCore","FLFishery","FLasher","FLBRP","mpb","FLife"), 
             repos="http://flr-project.org/R")
```

```{r}
library(FLCore)
library(FLasher)
library(FLBRP)
library(FLife)
```

### Devtools

The MyDas package has various methods for conditioning OMs, generating pseudo data and simulation testing data-limited stock assessment methods. As `mydas`  is still being developed, it is currently best installed from its [GitHub repository](https://github.com/lauriekell/mydas-pkg). It can be installed using the `devtools` package, (RTools)[https://cran.r-project.org/bin/windows/Rtools/] also needs to be installed, see (guide)[https://www.coursera.org/lecture/data-scientists-tools/installing-rtools-ecVeq].

```{r, eval=FALSE}
install.packages("devtools",dependencies=TRUE)
```

The `mydas` pakage can then be installed.

```{r, echo=TRUE, eval=FALSE}
library(devtools)

devtools::install_github("lauriekell/mydas-pkg")
```

```{r, echo=TRUE}
library(mydas) 
```

[Back to Top](#top)

# Example

## Operating Model

Create an Operating Model by specifying the life histoy parameters and filling in any missing values 

```{r, om-lh}
lh=FLPar(c(linf= 59.1,  k=0.28, t0=-0.4, a=0.01111,b=3.15,a50=4.0,l50=43.25),units="NA")

lh=lhPar(lh)   
```


Then create an `FLBRP` object that represents the equilibrium, i.e. expected dynamics and reference points.

```{r, om-eql, echo=TRUE}
eq=lhEql(lh)
```

Simulate a exploitation history and current depletion level.

The stock was originally underexploited, then effort increased until the stock was over fished at 2$F_{MSY}$. At which point a recovery plan was implemented, where F was reduced to $F_{MSY}$ in 1 generation time, after which the stock was exploited at $F_{MSY}$  

```{r, om-f, fig.height=3, fig.width=6, echo=TRUE}
gTime=c(round(gt(eq))) 

fbar(eq)=refpts(eq)["msy","harvest"]%*%FLQuant(c(rep(.1,19),
                                                 seq(.1,2,length.out=30),
                                                 seq(2,1.0,length.out=gTime)[-1],
                                                 rep(1.0,61)))[,1:105]
plot(fbar(eq))
```

**Figure `r iFig=iFig+1; iFig`** Exploitation History.


```{r}
om=as(eq,"FLStock")

om=fwd(om,f=fbar(om)[,-1], sr=eq)

plot(om)
```


## Catch per unit effort

An unbiased index of total biomass
```{r, survey, fig.height=4}
plot(stock(om))
```

**Figure `r iFig=iFig+1; iFig`** Stock biomass


```{r, cpue, fig.height=6}
plot(FLQuants(om,Survey=stock,
                 CPUE=function(x) catch(x)/fbar(x)))
```

**Figure `r iFig=iFig+1; iFig`** Comparison between unbiased survey and catch per unit effort

```{r, cpue2, fig.height=4, fig.width=6}
library(plyr)

dat=as.data.frame(FLQuants(om,Survey=stock,
                 CPUE=function(x) catch(x)/fbar(x)))
dat=ddply(dat,.(qname), transform, index=data/mean(data))
ggplot(dat)+geom_line(aes(year,index,col=qname))+
  theme(legend.position="bottom")
```

**Figure `r iFig=iFig+1; iFig`** Comparison between unbiased survey and catch per unit effort

### Effort creep

Model an increase in catchabilty, e.g. due to an increase in catchability.

```{r, trend, eval=TRUE}
trend<-function(object,bias=0.02) 
  object%*%FLQuant(cumprod(1+rep(bias,dim(object)[2])),dimnames=dimnames(object))
```


```{r}
plot(FLQuants(stock(om)%=%1,trend(stock(om)%=%1)))
```

**Figure `r iFig=iFig+1; iFig`** 2% increase per year in catchability.


```{r}
plot(FLQuants(catch(om)%/%fbar(om),trend(catch(om)%/%fbar(om))))
```

**Figure `r iFig=iFig+1; iFig`** Effect of a 2% increase per year in catchability on CPUE.

### Hyperstabilty

[Hyperstability](https://blog.nature.org/science/2016/05/11/hyperstability-the-achilles-heel-of-data-poor-fisheries/) is when catch rate stays stable while the actual fish population declines. 

```{r, hyperstability, eval=TRUE}
hyperstability<-function(object,omega=1,ref=apply(object,c(1,3:6),mean)) 
  ref%*%((object%/%ref)^omega)
```

```{r, hyper, fig.height=4}
om2=window(om,start=25,end=50)

plot(catch(om2)%/%fbar(om2))+
  geom_line(aes(year,data),
            data=as.data.frame(hyperstability(catch(om2)%/%fbar(om2),omega=0.5)),col="red")
```


**Figure `r iFig=iFig+1; iFig`** Effect of hyperstabilty (red), the stock appears to decline less. 


It is possible to develop indices tailored to different case studies

```{r plot,echo=TRUE, eval=TRUE}
plot(FLQuants(om,"Stock"   =stock,
                 "Adults"   =ssb,
                 "Recruits" =rec,
                 "Juveniles"=function(x) stock(x)-ssb(x),
                 "CPUE"     =function(x) catch(x)%/%fbar(x)))
```

**Figure `r iFig=iFig+1; iFig`** Different indices of abundance. 


# Uncertainty

```{r example2, fig.height=4, echo=TRUE, eval=TRUE}
cv=rlnorm(100,log(stock(om)),0.3)
plot(cv, iter=1)
```

**Figure `r iFig=iFig+1; iFig`** Measurement error, the red line is a single realisation. 


```{r example3, fig.height=4, echo=TRUE, eval=TRUE}
cv=rlnoise(100,log(stock(om)),0.3,0.8)
plot(cv, iter=1)
```

**Figure `r iFig=iFig+1; iFig`** Measurement error with auto-correlation, the red line is a single realisation. 



```{r example9,echo=TRUE,eval=TRUE}
set.seed(1234)

u =FLQuants("Unbiased"       =rlnorm(100,log(stock(om)),.3),
            "AR"             =rlnoise(100,log(stock(om)),.3,b=.7),
            "Hyperstability" =rlnorm(100,log(hyperstability(stock(om),0.5)),.3),
            "Trend"          =rlnorm(100,log(trend(stock(om),0.01)),.3),
            "Hetroscedascity"=rlnorm(100,log(stock(om)),.3)*trend(stock(om)%=%1,0.01),
            "Juvenile"       =rlnorm(100,log(stock(om)-ssb(om)),.3),
            "Mature"         =rlnorm(100,log(ssb(om)),.3),
            "Numbers"        =rlnorm(100,log(apply(stock.n(om),2:6,sum)),.3))

u=FLQuants(llply(u,function(x) x/mean(x)))
u=ldply(u,as.data.frame)

u.=ddply(u,.(year,.id), with, quantile(data,na.rm=T))
ggplot()+
  geom_line(aes(year,data,col=factor(iter)),
            data=subset(u,iter%in%c(2,11)))+
  geom_ribbon(aes(year,ymin=`25%`,ymax=`75%`),data=u.,col="grey",alpha=.5)+
  facet_wrap(~.id,ncol=2)+
  theme_bw()+theme(legend.position="none")
```


**Figure `r iFig=iFig+1; iFig`** Example of OEM for indices of abundance. 

### Mean length

Create an index of mean length
```{r}
mnLn=apply(wt2len(stock.wt(om),lh)%*%stock.n(om),c(2,6),sum)%/%apply(stock.n(om),c(2.,6),sum)
```

and compare with F
```{r}
plot(FLQuants(mnLn,fbar(om)))
```

**Figure `r iFig=iFig+1; iFig`** Comparison of mean length and fishing mortality. 


First create an Age Length Key that will be used to slice up numbers-at-age
```{r, turbot-alk,echo=TRUE,eval=TRUE}
ak=alk(lh)  
```

Then use the ALK to sample lengths
```{r, lfd}
lfd=lenSample(catch.n(om)[,20:65],ak,nsample=500)
```

```{r, turbot-oemplot,echo=TRUE,eval=TRUE}
ggplot(melt(lfd[,seq(1,45,10)]))+
  geom_histogram(aes(length,weight=value),binwidth=1)+
  facet_grid(year~iter,scale="free")+
  xlab("Length (cm)")+ylab("Frequency")+
  coord_cartesian(xlim=c(0,mean(lh["linf"])))
```

**Figure `r iFig=1; iFig`** Observation error model for turbot length samples. 

[Back to Top](#top)

[More information](#More information)

* You can submit bug reports, questions or suggestions on `FLPKG` at the `FLPKG` issue page ^[<https://github.com/flr/FLPKG/issues>], or on the *FLR* mailing list.
* Or send a pull request to <https://github.com/flr/FLPKG/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage ^[<http://flr-project.org>].
* The latest version of `FLPKG` can always be installed using the `devtools` package, by calling
```{r, devtools, echo=TRUE, eval=FALSE}
	library(devtools)
	install_github('flr/FLPKG')
```

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* FLPKG: `r # packageVersion('FLPKG')`
* **Compiled**: `r date()`
* **Git Hash**: `r system("git log --pretty=format:'%h' -n 1", intern=TRUE)`

## Author information

**Laurence Kell**. laurie@seaplusplus.es


## Acknowledgements

This vignette and many of the methods documented in it were developed under the MyDas project funded by the Irish exchequer and EMFF 2014-2020. The overall aim of MyDas is to develop and test a range of assessment models and methods to establish Maximum Sustainable Yield (MSY) reference points (or proxy MSY reference points) across the spectrum of data-limited stocks.

# References {#References}



[Back to Top](#top)
