---
title: "**FLife**"
subtitle: "Parallel Running"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
github_document:
  mathjax: TRUE
pdf_document:
  fig_width: 6 
  fig_height: 4 
tags: [FLR]
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
---

http://rpubs.com/pl202/436164

```{r knitr_init, echo=FALSE, results="hide"}
library(knitr)
## Global options
opts_chunk$set(cache     =TRUE,
               cache.path='cache/performance/',
               echo      =TRUE,
               eval      =TRUE,
               prompt    =FALSE,
               comment   =NA,
               message   =FALSE,
               warning   =FALSE,
               tidy      =FALSE,
               fig.height=6,
               fig.width =8,
               fig.path  ='tex/performance-')


iFig=0
```


```{r, pkgs, message=FALSE}
library(ggplotFL)
library(plyr)
library(reshape)
library(FLBRP)
library(FLife)
```

```{r, theme, echo=FALSE}
theme_set(theme_bw())
options(digits=3)
```

## Install FLR packages

To follow this tutorial you will have to install a number of packages, either from CRAN or from [www.flr-project.org](http://www.flr-project.org) where variety of packages and [tutorials](https://www.flr-project.org/doc/) are available.

Install FLR from [https://www.flr-project.org/](www.flr-project.org/)
```{r, eval=FALSE}
install.packages(c("FLCore","FLFishery","FLasher","FLBRP","mpb","FLife"), 
             repos="http://flr-project.org/R")
```

`devtools` needs to be installed and then loaded so that the `mydas` package can be installed from this GitHub repository.

```{r}
install.packages("devtools",dependencies=TRUE)
```

```{r, echo=TRUE, eval=FALSE}
library(devtools)

devtools::install_github("lauriekell/mydas-pkg")
```

## Load Libraries
```{r}
library(plyr)
library(reshape)
library(ggplot2)
```

```{r}
library(FLCore)
library(FLasher)
library(FLBRP)
library(FLife)
library(mydas)
```


```{r}
```



```{r}
library(FLCore)
library(FLasher)
library(FLBRP)
library(mydas)

library(doParallel)
library(foreach)

library(googledrive)

fls=drive_find()
drive_download("ray.RData",path="ray.RData")
load("ray.RData")

### Stochasticity
nits=dim(om)[6]
set.seed(1234)
srDev=FLife:::rlnoise(nits,FLQuant(0,dimnames=list(year=1:100)),0.2,b=0.0)

### OEM
uDev =FLife:::rlnoise(nits,FLQuant(0,dimnames=list(year=1:100)),0.3,b=0.0)

## MSE for Derivate empirical MP
scen=expand.grid(k1=seq(0.2,1.0,0.8),k2=seq(0.2,1.0,0.8),gamma=seq(1,2),
                 stringsAsFactors=FALSE)

registerDoParallel(4)
mseSBTD=mydas:::mseSBTD
empD<-foreach(i=(seq(dim(scen)[1])), 
              .combine=rbind,
              .multicombine=TRUE,
              .packages=c("FLCore","FLasher","FLBRP","FLife","plyr","reshape")) %dopar%{
                
               omRes=mseSBTD(om,eq,control=with(scen[i,],c(k1=k1,k2=k2,gamma=gamma)),start=60,end=100,
                             srDev=srDev,uDev=uDev)

               res  =cbind(scen=i,k1=scen[i,"k1"],k2=scen[i,"k2"],gamma=scen[i,"gamma"],
                            omSmry(omRes,eq,lh))
                
               res}
```

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* FLPKG: `r # packageVersion('FLPKG')`
* **Compiled**: `r date()`
* **Git Hash**: `r system("git log --pretty=format:'%h' -n 1", intern=TRUE)`

## Author information

**Laurence KELL**. laurie@seaplusplus.co.uk

## Acknowledgements


# References {#References}


