---
title: "**FLife**"
subtitle: "Example of Constructing an Operating Model"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
github_document:
  mathjax: TRUE
pdf_document:
  fig_width: 6 
  fig_height: 4 
tags: [FLR]
bibliography: refs.bib
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
---

```{r knitr_init, echo=FALSE, results="hide"}
library(knitr)
## Global options
opts_chunk$set(cache     =TRUE,
               cache.path='cache/turbot/',
               echo      =TRUE,
               eval      =TRUE,
               prompt    =FALSE,
               comment   =NA,
               message   =FALSE,
               warning   =FALSE,
               tidy      =FALSE,
               fig.height=6,
               fig.width =8,
               fig.path  ='tex/turbot-')


iFig=0
```


```{r, pkgs, message=FALSE}
library(ggplotFL)
library(plyr)
library(reshape)
library(FLBRP)
library(FLife)
```

```{r, theme, echo=FALSE}
theme_set(theme_bw())
options(digits=3)
```

## Install FLR packages

To follow this tutorial you will have to install a number of packages, either from CRAN or from [www.flr-project.org](http://www.flr-project.org) where variety of packages and [tutorials](https://www.flr-project.org/doc/) are available.

Install FLR from [https://www.flr-project.org/](www.flr-project.org/)
```{r, eval=FALSE}
install.packages(c("FLCore","FLFishery","FLasher","FLBRP","mpb","FLife"), 
             repos="http://flr-project.org/R")
```

`devtools` needs to be installed and then loaded so that the `mydas` package can be installed from this GitHub repository.

```{r}
install.packages("devtools",dependencies=TRUE)
```

```{r, echo=TRUE, eval=FALSE}
library(devtools)

devtools::install_github("lauriekell/mydas-pkg")
```

## Load Libraries
```{r}
library(plyr)
library(reshape)
library(ggplot2)
```

```{r}
library(FLCore)
library(FLasher)
library(FLBRP)
library(FLife)
library(mydas)
```

## Operating Model Conditioning using `FLife`

### Life history parameters
Get fishbase data
```{r}
load(url("https://github.com//fishnets//fishnets//blob//master//data//fishbase-web//fishbase-web.RData?raw=True"))
```

Select turbot as a case study
```{r}
lh=subset(fb,species=="Psetta maxima")

names(lh)[c(14,17)] = c("l50","a50")
lh=lh[,c("linf","k","t0","a","b","a50","l50")]

head(lh)
```

Get the means and create an `FLPar` object
```{r}
lh=apply(lh,2,mean,na.rm=T)
lh=FLPar(lh)
```

Values can be replace with any [better estimates](https://www.researchgate.net/publication/236650425_Ecological_and_economic_trade-offs_in_the_management_of_mixed_fisheries_A_case_study_of_spawning_closures_in_flatfish_fisheries) that may be available.

`lhPar` fills in missing values using life history theory, while quantities like selection-at-age are set using defaults, in this case set to be the same as maturity-at-age.  


```{r}
?lhPar
```

```{r}
par=lhPar(lh)
par
```

The parameters are then be used by `lhEql` to simulate the equilibrium dynamics by combining the spawner/yield per recruit relationships with a stock recruitment relationship, by creating an [`FLBRP` object](https://www.flr-project.org/doc/Reference_points_for_fisheries_management_with_FLBRP.html)

```{r}
eq=lhEql(par)

plot(eq)
```

To model time series the FLBRP object created by lhEql is coerced to an FLStock object and then [projected forward](https://www.flr-project.org/doc/Forecasting_on_the_Medium_Term_for_advice_using_FLasher.html)

For example a fishing time series is simulated that represents a stock that was originally lightly exploited, then effort is increased until the stock is overfished and then fishing pressure was reduced to recover the stock biomass to BMSY.

```{r}
fbar(eq)=refpts(eq)["msy","harvest"]%*%FLQuant(c(rep(.1,19),
                                              seq(.1,2,length.out = 30),
                                              seq(2,1.0,length.out = 10),
                                              rep(1.0,61)))[,1:105]
plot(fbar(eq))
```
Coercing the FLBRP object into an FLStock

```{r}
om=as(eq,"FLStock")
```
We can simply plot the forward F projection catch, recruitment and biomass estimates etc..
We call fwd() with the stock, the control object (fbar) and the stock recruitment relationship, and look at the results

```{r}
om=fwd(om,fbar=fbar(om)[,-1], sr=eq)
```
can change the units here also
```{r}
units(catch(om)) = units(discards(om)) = units(landings(om)) = units(stock(om)) = 'tonnes'
units(catch.n(om)) = units(discards.n(om)) = units(landings.n(om)) = units(stock.n(om)) = '1000'
units(catch.wt(om)) = units(discards.wt(om)) = units(landings.wt(om)) = units(stock.wt(om)) = 'kg'

plot(om)
```

### Stochasticity 
To add recruitment variability into add the residuals argument, e.g. for 100 Monte Carlo simulations with a CV of 0.3  
```{r}
nits=100

srDev=rlnoise(nits, rec(om) %=% 0, 0.3)

om=propagate(om,nits)

om=fwd(om,fbar=fbar(om)[,-1], sr=eq, residuals=srDev)
```

```{r}
plot(om)
```


**Figure `r iFig=iFig+1; iFig`** Simulate a stock with increasing F

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* FLPKG: `r # packageVersion('FLPKG')`
* **Compiled**: `r date()`
* **Git Hash**: `r system("git log --pretty=format:'%h' -n 1", intern=TRUE)`

## Author information

**Laurence KELL**. laurie@seaplusplus.co.uk

## Acknowledgements


# References {#References}


